Towards a standardized evaluation of imputation methodology

### 2022-07-22 Manuscript/Towards_a_standardized_evaluation_of_imputation_methodology.tex
add refs review morris

@@

When evaluating the statistical properties (and thereby the practical applicability) of imputation methodology, researchers most often make use of simulation studies. In such studies a complete dataset is usually generated from a statistical model, another model is used to induce missingness, and a set of evaluation criteria is postulated to evaluate the performance of one or more missing data methods. However, no gold standard has been established to evaluate imputation routines and, as a result, the validity and aim of the simulation setups may differ tremendously from one developer to another. Especially with novel imputation methods being propagated from the fields of machine learning and artificial intelligence, the differences may become more pronounced. Although these promising new methods seem to yield even sharper imputations than now-standard (semi-)parametric imputation methods, the comparison may not be fair due to the simulation setup.

The purpose of this paper is threefold: First, to raise some concerns with respect to evaluating imputation methodology. These concerns stem from careful consideration with fellow `imputers' and from encounters as a reviewer for statistical journals. Second, to provide imputation methodologists with a suggested course of action when simulating missing data problems. This suggested approach should identify common ground but is in no way intended as an absolute solution. This identifies the third purpose of this paper: discussion. We hope to elicit critical thinking regarding the problems at hand. We are all convinced that our methodology has some merit. But for sake of progress, it would be much more advantageous if the aim of our evaluations would go beyond \textit{proving the point} and would legitimately consider the statistical properties. [TODO: explain how the standardisation leads to more neutrality, but that this is just a set of elements to consider/report, not a complete 'cookbook'.]

@@

@@

>>>>

[TODO: add that if the goal is statistical inference, the uncertainty due to missingness should be quantified/captured by the missing data method. Refer to \citet{sper20} for difference in context of prediction. And paraphrase "the quality of a technique can only be evaluated with respect to the aims of the problem/s it's intended to solve in practice". we focus on comparative sim studies]

@@

When evaluating the statistical properties (and thereby the practical applicability) of imputation methodology, researchers most often make use of simulation studies. In such studies a complete dataset is usually generated from a statistical model, another model is used to induce missingness, and a set of evaluation criteria is postulated to evaluate the performance of one or more missing data methods. However, no gold standard has been established to evaluate imputation routines and, as a result, the validity and aim of the simulation setups may differ tremendously from one developer to another. Especially with novel imputation methods being propagated from the fields of machine learning and artificial intelligence, the differences may become more pronounced. Although these promising new methods seem to yield even sharper imputations than now-standard imputation methods, the comparison may not be fair due to the simulation setup.

The purpose of this paper is threefold: First, to raise some concerns with respect to evaluating imputation methodology. These concerns stem from careful consideration with fellow `imputers' and from encounters as a reviewer for statistical journals. Second, to provide imputation methodologists with a suggested course of action when using simulation studies to evaluate imputation techniques for missing data problems. This suggested approach should identify common ground but is in no way intended as an absolute solution. This identifies the third purpose of this paper: discussion. We hope to elicit critical thinking regarding the problems at hand. We are all convinced that our methodology has some merit. But for sake of progress, it would be much more advantageous if the aim of our evaluations would go beyond \textit{proving the point} and would legitimately consider the statistical properties. [TODO: explain how the standardisation leads to more neutrality, but that this is just a set of elements to consider/report, not a complete 'cookbook'.]

@@

[TODO: \citep{pawe22} pawel pre-reg simulation protocols? per-protocol results]

@@

[TODO: split up into design, execution and reporting? \citep{pawe22}]
